European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-24 manuscript/manuscript.qmd
Add para on broad picture of hegemony

The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power. European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts. American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds. This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.

>>>>

More broadly, the practice of IMF governance underscores that even the United States, as the system’s hegemon, operates under constraints. Securing outcomes typically requires consensus from other major powers, most notably European states, and compromises are often struck to maintain institutional cohesion. US influence also depends on a degree of legitimation, whether through the appearance of multilateral power-sharing, as in the transatlantic balance within the IMF, or through the promotion of ideological frameworks such as neoliberal economic orthodoxy. Yet such strategies can produce unintended consequences: Europe has not only embraced but also entrenched neoliberal doctrines, allowing it to leverage them in ways that enhance its own institutional authority. This dynamic illustrates a broader pattern in which American-founded institutions, originally designed to project US preferences, may evolve into arenas that constrain or even oppose US objectives, as exemplified by the UN Security Council’s opposition to the Iraq War. Hegemony, in this light, is both exercised and contested within the very structures the United States helped to create.
